Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-10 paper.tex
Some editing and extra text

Scientific applications differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}), implemented in consistent software 
libraries and running on clouds or dedicated commodity clusters with locality-aware 
file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications 
often involve binary data such as images and signals, processed by 
command-line tools using a mix of programming languages (C, Fortran, 
Python, shell scripts) and are deployed on large, shared clusters where 
data is fetched from data nodes to compute nodes using centralized shared file 
systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
applications and infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools. Moreover, performance is only one of the important
aspects to scientific computing. Other important aspects include reproducibility 
and provenance tracking. While these are minimally addressed by Big Data
processing engines, it is currently insufficient for scientific computing. 
Incorporation of provenance techniques into frameworks such are Spark are
currently being investigated~\cite{samba}.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(\texttt{tmpfs}) and memory-mapped files\todo{check that} might also 
explain the lack of adoption of Big Data engines for scientific 
applications.
In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Neuroimaging, our primary field of interest, is no exception to the 
generalized rise of data volumes in science, due to the joint increase 
of image resolution and subject cohort sizes~\cite{van2014human}. 
Processing engines have been developed with neuroinformatics 
applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
or PSOM~\cite{bellec2012pipeline}. Big Data engines have been used for 
neuroimaging applications too, including in the Thunder 
project~\cite{freeman2014mapping} and in more specific works such 
as~\cite{makkie2019fast}. However, no quantitative performance 
evaluation has been conducted on neuroimaging applications to assess the 
added-value of Big Data engines compared to traditional processing engines.

This paper addresses the following questions:
\begin{enumerate}
\item What is the effect of in-memory computing and data locality on neuroimaging applications?
\item Can in-memory computing be effectively enabled by the operating system rather than the data processing engine?
\end{enumerate}

Answers to these questions have important implications. 
In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
presented. It concludes that these systems need to be extended to 
better address scientific code integration, data partitioning, data 
formats and system tuning. We argue 
that such efforts should only be conducted if substantial performance 
improvements are expected from in-memory computing, lazy 
evaluation\todo{are we talking about lazy evaluation at all?} or data 
locality. On the other hand, scientific data processing engines are 
still being developed, and it is legitimate to wonder whether these 
projects should just be migrated to Spark, Dask, or other Big Data 
engines.

@@

@@

@@

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a LRU eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. It therefore becomes of great importance to persist to
memory and/or 
disk RDDs which will be reused throughout the application, particularly if the
recomputation is costly. 
%mention results in thunder paper

Data locality in Spark is achieved through the  scheduling of task to partitions which have the
data loaded in memory. If the data is instead stored on HDFS\todo{expand acronym if not done in intro}, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.
%improve paragraph

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
clusters. It is prepackaged with Hadoop installations. In contrast to YARN, 
Mesos was designed to be used in multitenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster. \todo{Maybe defer this to experiments}

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede Spark-cluster
scheduling time. Using pilot-scheduling strategies to add nodes to the Spark cluster
as they are allocated by the underlying schedulers may speedup allocation and 
overall processing time. This, however, is not studied in the current paper. \todo{Explain why}

@@

Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as SGE, PBS, SLURM and HTCondor\todo{expand acronyms}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype also includes many built-in 
interfaces to commonly used neuroimaging tools to be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable\todo{not the right word and sentence is difficult to parse.}
to existing Big Data solutions, which would necessitate modification to achieve this.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters. Data may reside in the engine memory, on 
a file system whose content reside in virtual memory (for instance 
\texttt{tmpfs}), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. The 
remainder of this Section explains this Table and provides related 
performance models.
\begin{table}
\centering
\begin{tabular}{c|ccc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality & Lazy \\
    \rowcolor{headcolor}
                                  & Computing     &               & Evaluation      \\
                                  \hline          
In-memory                         &   Y           & Y             & Spark               \\
\texttt{tmpfs}                    &   Y           & Y             & Spark           \\
Local Disk                        & Page Caching  & Y             & Spark           \\
Shared File System                & Page Caching  & N            & Spark          
% data locality is somehow present in shared fs due to page caching again.
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

@@

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in great 
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache. Such cache hits drastically speed-up data 
reads, by masking the disk latency and limited bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application 
\todo{check that}, which prevents processing engines to anticipate 
reads by preloading the cache. Scheduling strategies might be designed 
to maximize cache hits though. For instance, we hypothesize that lazy 
evaluation would result in more cache hits by scheduling data-dependent 
tasks on the same node.

Page caching has a more dramatic effect on disk writes, reducing their 
duration by several orders of magnitude. When a process calls the 
\texttt{write()} system call, data is copied to a memory cache that is 
asynchronously written to disk by flusher threads when memory shrinks, 
``dirty" (unwritten) data grows, or a 
process invokes the \texttt{sync()} system call. 
This asynchronous flushing of the page cache is called 
\emph{writeback}.
%~ It should be noted that writeback may lead to data 
%~ loss in case of system failure.

@@

This model applies to parallel applications assuming that (1) 
concurrent processes all write the same amount of data, (2) 
concurrent processes all consume the same CPU time, (3) data is written 
uniformly along task execution. Due these assumptions, all the 
processes will write at the same rate, which explains why the model 
does not depend on the total number of concurrent processes in the 
application, but only on the max number of concurrent processes 
executing on the same node ($\gamma$). We realize that these 
assumptions would usually be violated in practice, but this simple 
model already provides interesting insights on the performance of disk 
writes, as shown later. Naturally, the model also ignores other 
processes that might be writing to disk concurrently to the 
application, which we assume negligible here.

@@

% is there a page cahe when writing to a socket? there must be!
% anyway the model is in the form of Eq 1, but now delta is the network bandwidth
% and gamma is the total number of concurrent processes.
% also, is lustre going to slow down network writes?

%~ In any case, I think we want the following inequality to hold:
%~ ```
%~ D/C <= \Delta / \Gamma
%~ ```
%~ where Delta is now the apparent bandwidth of the file system and Gamma 
%~ is the max number of concurrent tasks in the cluster. Whether page 
%~ caching is used or not will affect what happens when the inequality 
%~ holds: if page caching is here, write time will be 0, otherwise it will 
%~ be non-zero. If the inequality doesn't hold then we have congestion and 
%~ it's a different game: transfer times might tend to infinity and data 
%~ loss is to be expected.

% Say that we don't use HDFS because we focus on multi-tenant clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).

% Make a table to explain where data locality and in-memory computing are
% available.  As a summary of the models.

% What happens when dataset and derived data is too large for disk?

In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.

@@

\subsection{Task Scheduling}

% Explain how we ensured that both engines lead to the same scheduling.
% Talk about pilot jobs and say we don't consider them here.

@@

All experiments were executed on the Dell Zenith cluster~\cite{dellHPC} running 
SLURM. For the 
Spark experiments, a Spark cluster was started atop our SLURM allocation. We
had access to 16 dedicated nodes. Each compute node ran on RHEL 7.4, was 
equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
network. The apparent bandwidth of the local disk was found to be 1480MB/S, 
whereas Lustre's bandwidth was found to be at 401.49MB/S.
% Dedicated Compute nodes with SLURM
% Lustre server

\begin{table}
\centering
\begin{tabular}{c|c}
$\delta/\gamma$ & $\Delta/\Gamma$\\
\hline
164.5~MB/s         & 3.2~MB/s
\end{tabular}
\label{table:infrastructure}
\caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
         maximum of 9 CPUs on each node.}
\end{table}

@@

% BigBrain
% Some BIDS dataset for fmriprep
For our experiments, BigBrain~\cite{bigbrain} a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old male's brain. The BigBrain was selected due 
to being the only one of its kind, as there does not yet exist a 
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing BigBrain has on the Big Data 
strategies, we partitioned the full 3845X3015X3470 voxel image into 30 (5X3X2)
chunks, 125 (5X5X5) chunks and 750 (5X15X10) chunks. Additionally, the full 
image was also split in half (769X603X347 voxels) and the half image was 
partitioned into 125 chunks.

@@

As it is possible that both small and large data may need to be processed using
the same pipeline \todo{explain pipelines in intro if not done already}, we 
wanted to determine the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to subject 1 of 
the ds000001 dataset version 6 ~\cite{openneuro}. In order to be able to split 
it into 125 chunks, it was necessary to zero-pad the image to the dimensions
165X200X200 voxels, which subsequently increased the total image size to 13M.

@@

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

For our first incrementation experiment, we investigated the effects of
in-memory computing on varying total data size. to do this, we increased the 
number of incrementation iterations from 1, 10 and 100 times. The total data 
size would then increased from 75GB, at 1 iteration, to 7500GB, at 100 
iterations. In order
to validate the model, the D/C was varied to fall in different regions of 
Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
and 1.9MB/s\todo{present this as a table?}. 80 and 1.9MB/s satisfied the 
inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
Lustre. These D/C ratios were applied to each iteration experiment, such that 
we were able to get both the effect of increasing the total amount of data, in 
addition to, increasing the task duration for a given data size.

We combined the first experiment with our second experiment such the we could 
evaluate the effects of in-memory computing on varying task duration. If the 
page cache has sufficient time to flush, it should be expected that using 
in-memory computing should have no increased speed-up over disk. We maintained 
the D/C ratios as those mentioned above, but rather than applying only one to
a different number of iterations, we applied all four D/C to all numbers of 
iterations. That is 1 iteration, 10 iterations and 100 iterations were all 
processed with using the aforementioned D/C ratios.

@@

\begin{table}
\centering
\begin{tabular}{c|c}
Application & $\frac{D}{C}$ (GB/s)\\
\hline
Big Brain            & [0-75] \\
fMRI pre-processing  & [0-1.5]
\end{tabular}
\label{table:applications}
\caption{Data-compute ratios for typical neuroinformatics applications}
\end{table}
% fMRI: about 100M in, 2GB out,2 30 min processing time to 10 hours

@@

% separate this in (1) the concepts, put in the model sections below, 
% (2) the implementation 'details', put in the experiments

% in-memory (Spark only)

% tmpfs: two extra in-memory copies. memfs could address that but we don't have it yet.
% Spark: when just passing filenames. turns out i need to set `preservePartitioning` to True
% local disk: write buffer issue

% shared fs: Lustre

@@

\subsection{Can \texttt{tmpfs} and Page Caches Emulate In-Memory Computing?}
% tmpfs and local disks fill up: need cleanup.
% Can we emulate in-memory computing using write buffers or tmpfs?

>>>>

Scientific applications differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. \todo{This sentence is extremely long} While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}), implemented in consistent software 
libraries and running on clouds or dedicated commodity clusters with 
locality-aware file systems such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), scientific applications often involve 
binary data such as images and signals, processed by command-line tools 
using a mix of programming languages (C, Fortran, Python, shell 
scripts), and are deployed on large, shared clusters where 
data is fetched from data nodes to compute nodes using centralized shared file 
systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
applications and infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools. Moreover, performance is only one of the important
aspects to scientific computing. Other important aspects include reproducibility 
and provenance tracking. While these are minimally addressed by Big Data
processing engines, it is currently insufficient for scientific computing. 
Incorporation of provenance techniques into frameworks such are Spark are
currently being investigated~\cite{samba}.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(tmpfs) and memory-mapped files\todo{check that} might also 
explain the lack of adoption of Big Data engines for scientific 
applications.
In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Neuroimaging, our primary field of interest, is no exception to the 
generalized rise of data volumes in science, due to the joint increase 
of image resolution and subject cohort sizes~\cite{van2014human}. 
Processing engines have been developed with neuroinformatics 
applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
or the Pipeline System for Octave and Matlab 
(PSOM~\cite{bellec2012pipeline}). Big Data engines have been used for 
neuroimaging applications too, including in the Thunder 
project~\cite{freeman2014mapping} and in more specific works such 
as~\cite{makkie2019fast}. However, no quantitative performance 
evaluation has been conducted on neuroimaging applications to assess the 
added-value of Big Data engines compared to traditional processing engines.

This paper addresses the following questions:
\begin{enumerate}
\item What is the effect of in-memory computing, lazy evaluation and data locality on current neuroimaging applications?
\item Can in-memory computing be effectively enabled by the operating system rather than the data processing engine?
\end{enumerate}

Answers to these questions have important implications. 
In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
presented. It concludes that these systems need to be extended to 
better address scientific code integration, data partitioning, data 
formats and system tuning. We argue 
that such efforts should only be conducted if substantial performance 
improvements are expected from in-memory computing, lazy 
evaluation\todo{are we talking about lazy evaluation at all?} or data 
locality. On the other hand, neuroimaging data processing engines are 
still being developed, and it is legitimate to wonder whether these 
projects should just be migrated to Spark, Dask, or other Big Data 
engines.

@@

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as XSEDE, Compute 
Canada or DEISA \todo{add links}, as this is the type of platform which 
is typically available to neuroscientists. We assume that HPC systems 
are multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available, for instance Lustre. We intentionally leave distributed, 
HDFS-like file system out, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{hadoop-on-demand} have not 
become mainstream yet.

@@

\todo{Maybe add an intro to this section, to outline it}

@@

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. It therefore becomes of great importance to persist to
memory and/or 
disk RDDs which will be reused throughout the application, particularly if the
recomputation is costly. 
%mention results in thunder paper

Data locality in Spark is achieved through the  scheduling of task to partitions which have the
data loaded in memory. If the data is instead stored on HDFS, the scheduler will
assign it to one of the preferred locations specified by HDFS. Spark's scheduler
utilizes delay scheduling to optimize fairness and locality for all tasks.
%improve paragraph

Three different types of schedulers are compatible with Spark: 1) Spark 
Standalone, 2) YARN~\cite{vavilapalli2013apache} and 3) Mesos~\cite{hindman2011mesos}. The Spark Standalone
scheduler is the default scheduler. YARN was a scheduler designed for Hadoop~\cite{white2012hadoop}
clusters. It is prepackaged with Hadoop installations. In contrast to YARN, 
Mesos was designed to be used in multitenant cluster environments. In our experiments,
we focus on the Spark Standalone cluster.

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede Spark-cluster
scheduling time. Using pilot-scheduling strategies to add nodes to the Spark cluster
as they are allocated by the underlying schedulers may speedup allocation and 
overall processing time. This, however, is not studied in the current paper.

@@

Nipype is a popular Python neuroimaging processing engine. It aims at being a 
solution for easily creating reproducible neuroimaging workflows. Although Nipype
does not employ any Big Data processing strategies, it provides access to numerous
plugins to schedulers found in most clusters readily available to researchers, 
such as SGE, PBS, SLURM and HTCondor\todo{expand acronyms}. It also includes its own scheduler, MultiProc,
for parallel processing on single nodes. Nipype also includes many built-in 
interfaces to commonly used neuroimaging tools to be incorporated within the 
workflows. Nipype's ability to easily parallelize workflows in researcher-available 
cluster setups, capture detailed provenance information necessary for reproducibility,
and allow users to easily integrate existing neuroimaging tools, make it preferable
to existing Big Data solutions, which would necessitate modification to achieve this.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose content reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         &   Y           & Y                       \\
tmpfs                             &   Y           & Y                   \\
Local Disk                        & Page Caching  & Y                   \\
Shared File System                & Page Caching  & N                
% data locality is somehow present in shared fs due to page caching again.
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 related to what happens when the processed data fills up the available 
 memory. In case engine memory is used, the engine has to cleanup 
 unused data with some strategy to avoid crashes. In case an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

@@

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in    
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
reads, by masking the disk latency and limited bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application 
\todo{check that}, which prevents processing engines to anticipate 
reads by preloading the cache. Scheduling strategies might be designed 
to maximize cache hits though. For instance, lazy 
evaluation could result in more cache hits by scheduling data-dependent 
tasks on the same node.

Page caching has a more dramatic effect on disk writes, reducing their 
duration by several orders of magnitude. When a process calls the 
\texttt{write()} system call, data is copied to a memory cache that is 
asynchronously written to disk by flusher threads, when memory shrinks, when
``dirty" (unwritten) data grows, or when a 
process invokes the \texttt{sync()} system call. 
This asynchronous flushing of the page cache is called 
\emph{writeback}.
%~ It should be noted that writeback may lead to data 
%~ loss in case of system failure.

@@

This model applies to parallel applications assuming that (1) 
concurrent processes all write the same amount of data, (2) 
concurrent processes all consume the same CPU time, (3) data is written 
uniformly along task execution. With these assumptions, all the 
processes will write at the same rate, which explains why the model 
does not depend on the total number of concurrent processes in the 
application, but only on the max number of concurrent processes 
executing on the same node ($\gamma$). We realize that these 
assumptions would usually be violated in practice, but this simple 
model already provides interesting insights on the performance of disk 
writes, as shown later. Naturally, the model also ignores other 
processes that might be writing to disk concurrently to the 
application, which we assume negligible here.

@@

% In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.

@@

@@

All experiments were executed on the Dell Zenith cluster~\cite{dellHPC} running 
SLURM. For the 
Spark experiments, a Spark cluster was started atop our SLURM allocation. We
had access to 16 dedicated nodes. Each compute node ran on RHEL 7.4, was 
equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
network. The apparent bandwidth of the local disk was found to be 1480MB/s, 
whereas Lustre's bandwidth was found to be at 401.49MB/s.
% Dedicated Compute nodes with SLURM
% Lustre server

%~ \begin{table}
%~ \centering
%~ \begin{tabular}{c|c}
%~ $\delta/\gamma$ & $\Delta/\Gamma$\\
%~ \hline
%~ 164.5~MB/s         & 3.2~MB/s
%~ \end{tabular}
%~ \label{table:infrastructure}
%~ \caption{Bandwidth thresholds on the target cluster using 15 worker nodes and a
         %~ maximum of 9 CPUs on each node.}
%~ \end{table}

@@

% BigBrain
% Some BIDS dataset for fmriprep
We used BigBrain~\cite{bigbrain}, a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old male's brain. The BigBrain was selected due 
to being the only one of its kind, as there does not yet exist a 
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing BigBrain has on the Big Data 
strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
image was also split in half ($769\times603\times347$ voxels) and the half image was 
partitioned into 125 chunks.

@@

As it is possible that both small and large data may need to be processed using
the same pipeline \todo{explain pipelines in intro if not done already}, we 
wanted to determine the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to subject 1 of 
the ds000001 dataset version 6 ~\cite{openneuro}. In order to be able to split 
it into 125 chunks, it was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

@@

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

\subsubsection{Setup}

For our first incrementation experiment, we investigated the effects of
in-memory computing on varying total data size. To do this, we increased the 
number of incrementation iterations from 1, 10 and 100 times. The total data 
size would then increase from 75GB, at 1 iteration, to 7500GB, at 100 
iterations. In order
to validate the model, the D/C was varied to fall in different regions of 
Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
and 1.9MB/s\todo{present this as a table? yes, I think it would be clearer (TG) but not mandatory.}. 80 and 1.9MB/s satisfied the 
inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
Lustre. These D/C ratios were applied to each iteration experiment, such that 
we were able to get both the effect of increasing the total amount of data, in 
addition to, increasing the task duration for a given data size.

We combined the first experiment with our second experiment such that we could 
evaluate the effects of in-memory computing on varying task duration. If the 
page cache has sufficient time to flush, it should be expected that using 
in-memory computing should have no increased speed-up over disk. We maintained 
the D/C ratios as those mentioned above, but rather than applying only one to
a different number of iterations, we applied all four D/C to all numbers of 
iterations. That is 1 iteration, 10 iterations and 100 iterations were all 
processed with using the aforementioned D/C ratios.

@@

@@

@@

\subsection{Can tmpfs and Page Caches Emulate In-Memory Computing?}
% tmpfs and local disks fill up: need cleanup.
% Can we emulate in-memory computing using write buffers or tmpfs?
